---
title: "Methods_hw4"
output: github_document
author: "Yishan Wang"
date: "2018-11-12"
---

# Problem 2

```{r include = FALSE}
library(tidyverse)
```

```{r}
heartdisease_data = read_csv("./data/HeartDisease.csv")
```

### a)

#### Description of the Data Set

The main outcome is `totalcost` of patients diagnosed with heart disease. The main predictor is `ERvisits`, which is number of emergency room visits. Other important covariates are `age`, `gender`, `complications` and `duration`. `interventions`, `drugs` and `comorbidities` are potential covariates.

#### Descriptive Statistics for all Variables of Interest

##### Descriptive statistics for continous variables of interest:

```{r}
heartdisease_data %>%
  select(totalcost, ERvisits, age, complications, duration) %>%
  summary()
```
?dplyr::select
##### Descriptive statistics for categorical variable of interest:

```{r}
table(factor(heartdisease_data$gender, levels = c(0, 1), labels = c('Male', 'Female'))) %>%
  addmargins()
```

### b)

#### Plot the distribution for variable `totalcost`:

```{r}
hist(heartdisease_data$totalcost, main = "Total Cost Distribution", xlab = "Total Cost ($)", col.main = "red", col.lab = "blue")
```

#### Use log transformation:

```{r}
hist(log(heartdisease_data$totalcost), main = "Total Cost Distribution", xlab = "Total Cost ($)", col.main = "red", col.lab = "blue")
```

### c)

#### Create a new variable called `comp_bin` by dichotomizing `complications`: 0 if no complications, and 1 otherwise.

```{r}
new_heartdisease_data = heartdisease_data %>%
  mutate(comp_bin = as.factor(ifelse(complications == 0, 0, 1))) %>%
  mutate(gender = as.factor(gender))
```

### d)

#### Fit a simple linear regression between the original `totalcost` and predictor `ERvisits`.

```{r}
ggplot(heartdisease_data, aes(x = ERvisits, y = totalcost)) +
  geom_point() +
  geom_smooth(method = 'lm',formula = y~x)

reg_original_slr = lm(totalcost ~ ERvisits, heartdisease_data)
summary(reg_original_slr)
```

##### Comments on significance and interpretation of the slope: 


#### Fit a simple linear regression between the transformed `totalcost` and predictor `ERvisits`.

```{r}
trans_heartdisease_data = heartdisease_data %>%
  filter(totalcost != 0) %>%
  mutate(trans_totalcost = log(totalcost)) 

ggplot(trans_heartdisease_data, aes(x = ERvisits, y = trans_totalcost)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = y~x)

reg_trans_slr = lm(trans_totalcost ~ ERvisits, trans_heartdisease_data)
summary(reg_trans_slr)
```

##### Comments on significance and interpretation of the slope: 

### c)

#### Fit a multiple linear regression with `comp_bin` and `ERvisits` as predictors.

```{r}
reg_original_mlr = lm(totalcost ~ ERvisits + comp_bin, new_heartdisease_data)
summary(reg_original_mlr)
```

##### I)

##### Test if `comp_bin` is an effect modifier of the relationship between `totalcost` and `ERvisits`. 

```{r}
reg_interaction = lm(totalcost ~ ERvisits + comp_bin + ERvisits * comp_bin, new_heartdisease_data)
summary(reg_interaction)
```

##### Comment

##### II)

##### Test if `comp_bin` is a confounder of the relationship between `totalcost` and `ERvisits`. 

```{r}

```

##### Comment

##### III)

### f)

##### I)

#####  Use the model in part e) and add additional covariates and fit MLR.

```{r}
full_model = lm(totalcost ~ ERvisits + comp_bin + age + gender + duration, new_heartdisease_data)
summary(full_model)
```

##### Comment

##### II)

```{r}
anova(reg_original_slr, full_model)
```

# Problem 3

```{r}
patsatisfaction_data = readxl::read_excel("./data/PatSatisfaction.xlsx") %>%
  janitor::clean_names()
```

### a)

#### Create a correlation matrix

```{r}
Hmisc::rcorr(as.matrix(patsatisfaction_data))
```

##### Initial Findings

### b)

##### Fit a multiple regression model and test whether there is a regression relation and test whether there is a regression relation.

```{r}
reg_mlr = lm(safisfaction ~ age + severity + anxiety, patsatisfaction_data)
summary(reg_mlr)
```

##### State the hypotheses, decision rule and conclusion.


### c)

```{r}
confint(reg_mlr, level = 0.95)
```

##### Interpret the coefficient and 95% CI associated with `severity`.

### d)

##### Obtain an interval estimate for a new patient’s satisfaction when Age = 35, Severity = 42, Anxiety = 2.1. 

```{r}
mean_Y_hat = function(dataset, x, y, n, conf.level = 0.95){

  reg = lm(y ~ x)
  
#Get the number of observations n
  n = dim(dataset)[1]
  
  Upper_limit = reg$coefficients[1] + reg$coefficients[2]*x +
    qt(0.5*(1 + conf.level), n - 2) * sqrt(sigma(reg)^2*(1/n + (x - mean(x))^2/sum((x - mean(x))^2)))
  Lower_limit = reg$coefficients[1] + reg$coefficients[2]*x -
    qt(0.5*(1 + conf.level), n - 2) * sqrt(sigma(reg)^2*(1/n + (x - mean(x))^2/sum((x - mean(x))^2)))
  list(Upper_limit, LL = Lower_limit)
}
```

?dim

