---
title: "Methods_hw4"
output: github_document
author: "Yishan Wang"
date: "2018-11-12"
---

# Problem 2

```{r include = FALSE}
library(tidyverse)
library(Hmisc)
```

```{r}
heartdisease_data = read_csv("./data/HeartDisease.csv")
```

### a)

#### Description of the Data Set

The main outcome is `totalcost` of patients diagnosed with heart disease. The main predictor is `ERvisits`, which is number of emergency room visits. Other important covariates are `age`, `gender`, `complications` and `duration`. `interventions`, `drugs` and `comorbidities` are potential covariates.

#### Descriptive Statistics for all Variables of Interest

##### Descriptive statistics for continous variables of interest:

```{r}
heartdisease_data %>%
  select(totalcost, ERvisits, age, complications, duration) %>%
  summary()
```

##### Descriptive statistics for categorical variable of interest:

```{r}
table(factor(heartdisease_data$gender, levels = c(0,1), labels = c('Male','Female'))) %>%
  addmargins()
```

### b)

#### Plot the distribution for variable `totalcost`:

```{r}
hist(heartdisease_data$totalcost, main = "Total Cost Distribution", xlab = "Total Cost ($)", col.main = "red", col.lab = "blue")
```

#### Use log transformation:

```{r}
hist(log(heartdisease_data$totalcost), main = "Total Cost Distribution", xlab = "Total Cost ($)", col.main = "red", col.lab = "blue")
```

### c)

#### Create a new variable called `comp_bin` by dichotomizing `complications`: 0 if no complications, and 1 otherwise.

```{r}
new_heartdisease_data = heartdisease_data %>%
  mutate(comp_bin = ifelse(complications == 0, 0, 1))
```

### d)

#### Fit a simple linear regression between the original `totalcost` and predictor `ERvisits`.

```{r}
ggplot(heartdisease_data, aes(x = ERvisits, y = totalcost)) +
  geom_point() +
  geom_smooth(method = 'lm',formula = y~x)

reg_original = lm(totalcost ~ ERvisits, heartdisease_data)
summary(reg_original)
```

##### Comments on significance and interpretation of the slope: 


#### Fit a simple linear regression between the transformed `totalcost` and predictor `ERvisits`.

```{r}
trans_heartdisease_data = heartdisease_data %>%
  filter(totalcost != 0) %>%
  mutate(trans_totalcost = log(totalcost)) 

ggplot(trans_heartdisease_data, aes(x = ERvisits, y = trans_totalcost)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = y~x)

reg_trans = lm(trans_totalcost ~ ERvisits, trans_heartdisease_data)
summary(reg_trans)
```

##### Comments on significance and interpretation of the slope: 

